---
title: "Levelized cost of energy analysis"
output:
  flexdashboard::flex_dashboard:
    orientation: rows
    vertical_layout: scroll
    mathjax: null
runtime: shiny
---

```{r setup, include=FALSE}
#knitr::opts_knit$set(root.dir = "..")
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, error = FALSE)
```

```{r, echo = FALSE}
suppressMessages({
  #library(flexdashboard)
  library(plotly)
  library(ggplot2)   # remove?
  library(DT)
  # library(data.table)
  # library(reshape2)
  library(FinancialMath)
})

downloadButtonRmd <- function (outputId, label = "Download", class = NULL, ...)  {
     tags$a(id = outputId, class = paste("btn btn-default shiny-download-link", 
        class), href = "", target = "_blank", download = NA, 
        icon("download"), label, ...)
 }
```

    
Inputs {.sidebar data-width=300}
-----------------------------------------------------------------------

```{r}
#actionButton("calculate", "Calculate")

numericInput("capacity_mwh", "Capacity (MW)", 175)
sliderInput("capacity_pct", "Capacity factor (%)", 0, 100, value = 38)
numericInput("fixed_om", "Fixed O&M ($/kW-year)", 36)
numericInput("variable_om", "Variable O&M ($/MWh)", 0)

## Capital structure
sliderInput("debt", "Debt (as % of investment)", 0, 100, value = 60)
numericInput("cost_debt", "Cost of debt (%)", 8)
numericInput("cost_equity", "Cost of equity (%)", 12)

# Taxes and depreciations
sliderInput("tax_rate", "Tax Rate (%)", 0, 100, value = 40)
numericInput("economic_life", "Economic Life (years)", 20)

## Capex
numericInput("capex", "Investment ($/kW)", 1350)

downloadButtonRmd('report','Download')
```

```{r}
#calculate <- eventReactive(input$calculate, {
calculate <- reactive({
  year <- 1:round(input$economic_life)
  generation <- rep(365 * 24 * input$capacity_mwh * input$capacity_pct / 100, length(year))
  
  total_om <- generation * input$variable_om + 1000 * input$capacity_mwh * input$fixed_om
  
  # debt 
  total_debt   <- 1000 * input$capex * input$capacity_mwh * input$debt / 100
  total_equity <- 1000 * input$capex * input$capacity_mwh * (1 - input$debt / 100)
  tmp <- amort.table(total_debt, n = length(year), pmt = NA, i = input$cost_debt / 100, ic = 1, pf = 1, plot = FALSE)
  tmp <- data.frame(tmp$Schedule)
  colnames(tmp) <- c("payment", "interest", "principal", "debt")
  
  debt <- tmp$debt
  interest <- tmp$interest
  principal <- tmp$principal
  debt_serviced <- tmp$payment
  
  depreciation <- rep(total_debt / max(year), length(year))   # linear!
  
  # function to calculate the energy cost
  
  npv <- function(price) {
    revenues <- generation * price
    ebitda <- revenues - total_om
    
    taxable <- ebitda - depreciation - interest
    tax <- taxable * input$tax_rate / 100
    
    net_cash_flow <- ebitda - debt_serviced - tax
    
    iir <- sum(net_cash_flow / (1 + input$cost_equity / 100)) - total_equity
  }
  
  my_price <- uniroot(npv, interval = c(0, 10000))$root
  
  price <- rep(my_price, length(year))
  revenues <- generation * my_price
  ebitda <- revenues - total_om
  
  taxable <- ebitda - depreciation - interest
  tax <- taxable * input$tax_rate / 100
  
  net_cash_flow <- ebitda - debt_serviced - tax
  
  out <- data.frame(
    year = year,
    generation = generation,
    price = my_price,
    revenues = revenues,
    total_om = total_om,
    ebitda = ebitda,
    debt = debt,
    interest = interest,
    principal = principal,
    serviced = debt_serviced,
    depreciation = depreciation,
    taxable = taxable,
    tax = tax,
    net_income = net_cash_flow
  )
  
})
```


Row
-------------------------------------

### Levelized cost of energy

```{r}
renderDataTable({
  
  out <- calculate()
  
  out$generation <- round(out$generation / 1000, digits = 1)
  out$price      <- round(out$price, digits = 1)
  out$revenues   <- round(out$revenues / 1e6, digits = 1)
  out$total_om   <- round(out$total_om / 1e6, digits = 1)
  out$ebitda     <- round(out$ebitda / 1e6, digits = 1)
  out$debt       <- round(out$debt / 1e6, digits = 1)
  out$interest       <- round(out$interest / 1e6, digits = 1)
  out$principal       <- round(out$principal / 1e6, digits = 1)
  out$serviced       <- round(out$serviced / 1e6, digits = 1)
  out$depreciation       <- round(out$depreciation / 1e6, digits = 1)
  out$taxable       <- round(out$taxable / 1e6, digits = 1)
  out$tax       <- round(out$tax / 1e6, digits = 1)
  out$net_income       <- round(out$net_income / 1e6, digits = 1)
  

  datatable(out,
        rownames = FALSE,
        colnames = c("Year", "GWh/year", "$/MWh", "Revenue", "O&M", "EBITDA", "Debt remaining", "Interest Payment", "Principal Payment", 
                     "Debt Payment", "Depreciation", "Taxable Income", "Tax", "Net Income"),
        options = list(
          pageLength = max(out$year),
          paging = TRUE,
          searching = TRUE,
          ordering = FALSE,
          dom = "t")) #%>% 
    # formatRound("presupuesto_sin_impuesto", 3) %>% 
    # formatPercentage("baja_media", 1) %>% 
    # formatPercentage("baja_temeraria", 1) %>% 
    # formatRound("presupuesto_sin_impuesto", 2)
})
```

**Note:** Unless explictly stated otherwise, amounts are expressed in millions of USD.

```{r}
output$report <-downloadHandler(
  filename = function() "levelized_cost_energy.csv",
  content = function(file) {
        write.csv(calculate(), file = file, row.names = FALSE)
      },
  contentType="text/csv"
)
```


Row
-------------------------------------

### Notes

This dashboard is a reverse engineered, dynamic, and slightly simplified version of the spreadsheet embedded in [Lazard's Levelized Cost of Energy Analysis (v. 15.0)](https://www.lazard.com/media/451905/lazards-levelized-cost-of-energy-version-150-vf.pdf).

By default, it shows the analysis for the wind farm shown in page 14 of such document. It is possible to reproduce the levelized costs of energy of the plants/scenarios shown in pages 16-19.

The simplifications in this dashboard (with respect to Lazard's spreadsheet) are:

* O&M costs do not increase overtime.
* The discounting method used here is simpler (constant discount throughout the plant's life).
* The cost of fuel, if any, needs to be explicitly calculated and inputed as a variable O&M cost.


